Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching

### 2025-09-09 sections/setup_and_posthoc_analysis.tex
Updates from Overleaf

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\DeclareMathOperator*{\argmin}{arg\,min}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{makecell}
% my added packages and commands =======================
\usepackage[numbers]{natbib} %removed cite package
\usepackage{tikz}
\usetikzlibrary{positioning,shapes.geometric}
\usepackage{adjustbox}
\usepackage{listings}
\usepackage{caption}

\newcommand{\highlight}[1]{\textcolor[RGB]{00,100,100}{#1}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

\captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
\providecommand{\gls}[1]{#1}
\definecolor{rank1}{HTML}{70FF70}
\definecolor{rank2}{HTML}{858585}
\definecolor{rank3}{HTML}{454545}
\definecolor{rank4}{HTML}{000000}
\newcommand{\SIMSESpec}{\texttt{SIMSE\_Spec}}
\newcommand{\LoneSpec}{\texttt{L1\_Spec}}
\newcommand{\JTFS}{\texttt{JTFS}}
\newcommand{\DTWEnv}{\texttt{DTW\_Envelope}}
\newcommand{\LossSelect}{\textbf{Loss Selection}}
\newcommand{\SynthSelect}{\textbf{Synthesis Selection}}
\newcommand{\PeriodicLoss}{\textbf{Periodic Loss}}
\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
\newcommand{\BPNoise}{\textbf{BP-Noise}}  
\newcommand{\AddSineSaw}{\textbf{Add-SineSaw}}  
\newcommand{\AmpMod}{\textbf{Noise-AM}}  
\newcommand{\FMMod}{\textbf{SineSaw-AM}}

\input{code_style}

%  ================================
\begin{document}
\title{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}

\author{Amir Salimi, Abram Hindle, Osmar R. Za{\"i}ane}

% \markboth{IEEE TRANSACTIONS ON AUDIO, SPEECH AND LANGUAGE PROCESSING, VOL -, 2025}%
% {Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}

\markboth{%
  \parbox{\textwidth}{%
    This work has been submitted to the IEEE for possible publication. \\
    Copyright may be transferred without notice, after which this version may no longer be accessible.%
  }%
}{}

\maketitle

\begin{abstract}
Manual sound design with a synthesizer is inherently iterative: an artist compares the synthesized output to a mental target, adjusts parameters, and repeats until satisfied. Iterative sound-matching automates this workflow by continually programming a synthesizer under the guidance of a loss function (or similarity measure) towards a target sound. Prior comparisons of loss functions have typically favored one metric over another, but only within narrow settings: limited synthesis methods, few loss types, often without blind listening tests. This leaves open the question of whether a universally optimal loss exists, or the choice of loss remains a creative decision conditioned on the synthesis method and the sound designer's preference. We propose differentiable iterative sound-matching as the natural extension of the available literature, since it combines the manual approach to sound design with modern advances in machine learning. To analyze the variability of loss function performance across synthesizers, we implemented a mix of four novel and established differentiable loss functions, and paired them with differentiable subtractive, additive, and AM synthesizers. For each of the sixteen synthesizer–loss combinations, we ran 300 randomized sound-matching trials. Performance was measured using parameter differences, spectrogram-distance metrics, and manually assigned listening scores. We observed a moderate level of consistency among the three performance measures. Our post-hoc analysis shows that the loss function performance is highly dependent on the synthesizer. These findings underscore the value of expanding the scope of sound-matching experiments, and developing new similarity metrics tailored to specific synthesis techniques, rather than pursuing one-size-fits-all solutions.
\end{abstract}

% the automatic performance measures have not been compared to human hearing
% lack of creativity in loss function definitions 
\begin{IEEEkeywords}
Audio synthesis, differentiable digital signal processing, music information retrieval, sound-matching
\end{IEEEkeywords}

% need to simplify the intro, and expand the hypothesis
\section{Introduction}
Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer~\cite{creativecherep2024}. Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation. However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds. One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods. As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.

A digital audio synthesizer is any software used for the creation and manipulation of audio. Since the advent of digital signal processing (\gls{DSP}) in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}. Manual sound design with a synthesizer is an iterative process of trial and error; when the artist has a desired (or target) sound in mind, sound design involves listening to the output, determining whether the output is closer or farther from the target, and adjusting the parameters of the synthesizer accordingly until the target is reached~\cite{russ1999sound}. This paper explores the automation of this process, which we call \textit{iterative sound-matching}.

Automatic sound design has been a subject of research for decades, with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}. \textit{Sound-matching} is a common form of automatic sound design explored in literature, where a similarity metric (or loss function) is given, and the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}. Our analysis of past works shows that this general goal of ``sound-matching'' can be broken down in to very different subtypes, many of them rarely or never explored, despite their relevance to practical applications. This leads us to a typology of sound-matching tasks, and the importance of \textit{differentiable, iterative sound-matching}, rarely explored in previous literature~\cite{vahidi2023mesostructures}.

\textit{Iterative sound-matching} refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met. The two necessary components are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters. In the iterative workflow, a target sound is selected, the synthesizer outputs sounds based on its parameters, and the loss function determines the proximity of the output to the target. From there, the synthesizer parameters are updated, the output and target are compared, and so on.

Despite the early popularity of the iterative approach~\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}, recent works in sound-matching have favored supervised learning using inference models, deep neural-networks (\gls{DNNs}), and large datasets. Rather than continually interacting with the environment, these models learn from a dataset of inputs (synthesizer parameters) and outputs (corresponding sound), and predict the synthesizer parameters for new sounds~\cite{yee2018automatic,han2023perceptual,han2024learning,masuda2021soundmatch,masuda2023improving}.

Iterative sound-matching is interesting for multiple reasons. First, it better resembles the human approach to sound-design, which requires continually interacting with the environment and making decisions according to a feedback signal~\cite{sutton2018reinforcement}. Second, the trial-and-error approach leaves a trail of parameter sets from the starting point to the end goal, which could be of use to sound designers. Third, relative to the DNN-based inference models, it provides a more tractable environment with room for creative approaches to sound-similarity definitions, and more amenable to deployment on different synthesizers as it does not require pre-training on large datasets.

This work takes a \textit{differentiable} approach to iterative sound-matching, a natural extension of early works in iterative sound-matching combined with recent advances in Machine Learning (\gls{ML}) software~\cite{braun2024dac,paszke2019pytorch} and increase in computational power~\cite{moore2012introduction,schaller1997moore}. A differentiable approach requires differentiable loss and synthesizer functions, allowing the synthesizer parameters to be updated using gradient descent~\cite{goodfellow2016deep,boyd2004convex}.

We discuss the history of sound-matching, and provide a nomenclature for different subtypes of sound-matching tasks. A number of weaknesses in current literature (see Section~\ref{sec:lacking}) are identified. The experiments in this work focus on two of these core issues, referred to as \LossSelect~and \SynthSelect. The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.

In the case of \LossSelect, there have been many works and experiments comparing the accuracy of loss functions and choosing the best performers~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}. We note that these claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings. One such limitation is the synthesizers used, and ties into the problem of \SynthSelect, as typically only a single (and often, simple) synthesizer has been used for testing the effectiveness of loss functions. Furthermore, there is a lack of diversity in the underlying functions used by the synthesizers, with simple Frequency Modulation (\gls{FM}) being the most common method of synthesis~\cite{horner1993machine,vahidi2023mesostructures,mitchell2007evolutionary,caspe2023envelope}.

Considering the interplay between the two problems, we design experiments that address them simultaneously. Testing the hypothesis that the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis. We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance as measured by the similarity of the final output to the target. We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors. We compare the distributions of the evaluation scores in order to rank the loss functions from best to worst. The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching.

We note that for some synthesizers, our novel use of ``dynamic time warping of envelopes'' as a loss function for differentiable sound-matching outperforms the standard methods of audio comparison. These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods. Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.

\section{Background}
\label{sec:background}
In this Section, we give a formal definition of sound-matching and the necessary components. These discussions lead to a nomenclature for types of sound-matching, as well as major perennial issues in the field.

\subsection{Digital Audio Synthesis}
\label{sec:diffSynth}
A digital audio synthesizer is any software used for the creation and manipulation of digital audio. Digital synthesizers use signal processing chains with a variety of digital signal processing (\gls{DSP}) functions to create sounds. These functions are often parametric, and the set of parameters given to the synthesizer's chain of DSP functions is called a synthesizer program.

The simplest form of DSP function is a sinusoidal tone. For example, consider:
\[ x(n) = sin( 2\pi n T)\]

where $T$ represents time in seconds, and $n$ is the frequency parameter. If $n$ is 1, then $x(n)$ would be a waveform with a frequency of 1 hertz, meaning that the value of $x(n)$ oscillates to the same point once every second. Waveform generators are called oscillators for this reason.

% If $n$ is in the range of 30 to 20,000 (the upper limit is often lower for some), $x(n)$ could be played as an audible tone~\cite{smith2007mathematics,jilek2014reference}.

Since the advent of digital signal processing in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}. This includes oscillators, filters, equalizers, and envelopes, which can be used sequentially or in parallel~\cite{lyons1997understanding,russ1999sound}. Sound design with a synthesizer is done by modifying the parameters of the functions until reaching a desired output~\cite{roads1996computer,pinch2004analog}.

\todo{A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}. }

Differential DSP (\gls{DDSP}) is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP~\cite{engel2020ddsp}. Implementing complex DSP functions in a differentiable manner can be challenging, and effective differentiable audio similarity measures require careful mathematical expression of the desired attributes of sound. These issues have likely contributed to the limited exploration of this domain. Perhaps the most relevant recent work is by Vahidi \textit{et al.}, featuring the comparison of two differentiable loss functions with a simple FM synthesizer~\cite{vahidi2023mesostructures}.

\subsection{Sound Representation and Loss Functions}
\label{sec:loss_funcs}
A digital sound (or an audio signal) is a series of numbers~\cite{smith1991viewpoints,smith2007mathematics}. To compare two digital sounds, the two corresponding series are passed to a function that measures their similarity. Two signals can sound identical to our ears, without having any values in common~\cite{moore2012introduction}. This necessitates the use of proxy representations (or feature extractors) when comparing sounds automatically. Similarity between the target sound and the synthesizer output is then measured by some form of subtraction and summation of the proxy representations.

In sound-matching, particularly in a Deep Learning (\gls{DL}) context~\cite{goodfellow2016deep}, the similarity function can also be called a \textit{loss} function, where the emphasis is on measurement and reduction of the distance between target and output. It is important to note that there is a close relationship between the loss function $L$ and the sound representation function $\phi$. $L$ is the result of a distance measure $d$ applied to the features extracted by $\phi$.

\[
L(\theta, t) = d\langle\phi(t),\phi(x)\rangle
\]

\noindent

A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry,richard2025model}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
\[
L(\theta, t) = \left\| \phi(t) - \phi(x) \right\|_1
\]

Here we discuss four methods of audio representation and the corresponding loss functions. Three of which have been used in previous works, and a novel method not previously used in iterative sound-matching.

\subsubsection{Parameter Loss}
A common measure of similarity in sound-matching is the distance between synthesizer parameter sets, referred to as ``P-Loss"~\cite{han2023perceptual}. Typically, for the implementation of P-Loss the parameter sets are treated as vectors in space, and L1 or L2 distance is applied. There are two major limitations to this approach: First, the target and output sound must be made by the same synthesizer; otherwise the parameter sets cannot be compared (see Section~\ref{sec:matching_types}). Second, the relationship between synthesizer parameters and the audio output is not linear~\cite{shier2020spiegelib,han2023perceptual,esling2019flow}.

\subsubsection{Fourier Spectrograms}
\label{sec:fourier_specs}
Fourier-based transformations such as short-time Fourier transforms (\gls{STFT}), Mel-spectrograms, and Mel-frequency cepstral coefficients have been viewed as the de facto and state-of-the-art representation of audio~\cite{beauchamp2003error,mitchell2007evolutionary,yee2018automatic}, however, there are many issues associated with their use in sound-matching~\cite{turian2020sorry,vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}. Fourier transformations allow for the conversion of a signal from the time-domain to the frequency domain. Audio spectrograms can be generated by segmentation of a piece of audio into overlapping windows followed by the application of Fourier transforms to each window. They are costly to compute, but provide a better temporal view of changes in frequency content~\cite{muller2007dynamic,smith2007mathematics}. There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.

What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach. For example, Mel-Spectrograms \textit{bin} frequencies on a near-logarithmic scale to better match human perception of frequencies~\cite{muller2007dynamic}. Multi-scale spectrograms (\gls{MSS}) used in recent works are a simple weighted average of multiple spectrograms with different parameters such as window size, number of frequency bins, and hop size~\cite{engel2020ddsp,vahidi2023mesostructures}; this may provide some improvements at a higher computational cost~\cite{turian2020sorry,engel2020ddsp}.

\subsubsection{Joint-Time Frequency Spectrum}
Recent works have focused on the limitations of parameter and spectral loss functions in sound-matching~\cite{vahidi2023mesostructures,uzrad2024diffmoog}, seeking to create more effective general solutions for the comparison of audio. 
Noting the aforementioned weaknesses of comparing STFT spectrograms, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}. JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint}.

\subsubsection{Dynamic Envelope Warping}
Dynamic Time Warping (DTW) is a method for measuring similarity between multi-dimensional time-series~\cite{rabiner1993fundamentals,muller2007dynamic,giorgino2009computing}. Given any two time-series $X = \{x_1,x_2,...,x_m\}$ and $Y = \{y_1,y_2,...,y_n\}$, we have indices $i\in\{1...m\}$ and $j\in\{1...n\}$ defining $X$ and $Y$. When the series are \textit{warped}, these indices change to expand or contract different portions of the series. To borrow the notation given by Muller~\cite{muller2007dynamic}, warped indices are a sequence $p=(p_1,...,p_L)$, where \(p_\ell = (m_\ell, n_\ell) \in [1 : m] \times [1 : n] \text{ for } \ell \in [1 : L]\), meaning that the indices for $X$ and $Y$ are reorganized under special conditions. In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}.

DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance. The distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}. Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.
% \textcolor{red}{DTW was originally used for signal comparison but using it the way we do is novel}

\subsection{Automatic Sound Matching}
\label{sec:sound_matching_definition}
The problem of sound-matching has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$. To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}. The major components are as follows: 
\begin{itemize}
    \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
    \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
    \item $t$: The target sound which we want to replicate or imitate. 
    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
    \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
\end{itemize}

\highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}. Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.

\subsection{Sound Matching Types}
\label{sec:matching_types}

Sound-matching tasks can be categorized based on the choices of target sound $t$, the loss $L$ and representation function $\phi$, and the heuristic used for minimizing $\theta$. We will discuss the subtypes of sound-matching tasks in the upcoming sub-sections.

\subsubsection{In-Domain vs Out-of-Domain}
\label{sec:in-domain}
The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, which we call \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}. To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $t$, or put differently, if $t$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}. If $t$ is not an output of $g$, then the sound-matching task is \textit{out-of-domain}. In-domain tasks in general are simpler, and often there is a guarantee that there is a correct answer to the sound-matching problem, particularly if the goal is accurate replication of the sound. If the target sound is out-of-domain, replication is not guaranteed, and the goal becomes the \textit{imitation} of some aspect of sound.

\subsubsection{Replication vs Imitation}
Regardless of the domain, the generation goal can be \textit{replication} or \textit{imitation} of the target sound. In replication, the goal is to make an identical copy of the target sound. Imitation is an artistic pursuit and harder to define, since the goal is to make new sounds that only retain a subset of the target's sonic features. While closely related to the in-domain versus out-of-domain problem, the choice of replication versus imitation is more dependent on how the loss and representation functions are defined.

% From a signal perspective, replication is a scenario where the goal is to find a $\theta$ such that $t-x=0$ where $x=g(\theta)$. From a sound perspective, this definition of replication is unnecessarily restrictive, as $t$ and $x$ can sound identical but contain very different values. For example, if $t$ and $x$ were periodic waveforms at an audible frequency $f$ with different phase values, their subtraction would not be 0, yet they would sound identical to our ears, as our ears are insensitive to phase values~\cite{moore2012introduction}. Representation functions such as STFT spectrograms or Mel-frequency cepstral coefficients (\gls{MFCC}) become viable in these cases, as they allow for separation and removal of the phase information within an audio signal. The \textit{sound replication} goal can be defined, in layman's terms, as ``$x$ sounds exactly like $t$". More formally, the goal is to find parameters such that $\phi(t)-\phi(x)=0$, where $\phi$ transforms an audio into a spectrogram that removes the phase information from sounds, but otherwise represents the sonic information as accurately as possible.

% With \textit{sound imitation}, the goal is the replication of some aspect of the target sound. Instrument conversion is a common example of imitation. As an example, a note played by a violin is converted to the same note played by a piano. Conversion of even longer pieces has been made possible with large enough datasets~\cite{engel2020ddsp}. Imitation of sound envelopes (how the loudness changes over time) is another interesting idea~\cite{caspe2023envelope}. Transfer of other sound properties such as reverb and delay has also been a subject of research~\cite{engel2020ddsp}.

% A target sound can have many properties of interest, and the imitation task is the transfer of a subset of those properties to other sounds. Relative to replication, imitation is a very open ended goal. From a research point of view, success in the replication task can be measured more reliably, often without the need for manual evaluation. With imitation, listening tests are needed to ensure quality of outputs, particularly in a supervised learning setting where learning is done by examples~\cite{salimi2020make,masuda2021soundmatch}. The need for subjective evaluations has perhaps contributed to the lack of research in imitation and out-of-domain sound-matching.

\subsubsection{Supervised Versus Direct Optimization}
\label{sec:optimization}

The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound. 
\[
\theta^* = \arg\min_{\theta} L(\theta,t)
\]

The heuristics in past works can be broadly split into two categories:
\begin{enumerate}
    \item Supervised (or inference) methods which use datasets of sound and parameter pairs to describe the generation objective, often with a large number of input and output examples~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}
    \item Unsupervised or \textit{direct sound optimization} methods~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}. Usually require at most one target example, and a loss function that measures the difference between the target and synthesizer outputs. 
\end{enumerate}

\todo{
Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}. These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values. The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.}

 Differentiable settings have the benefit of making goal-oriented updates to the synthesizer parameters (the goal being the minimization of loss). There are also drawbacks: other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, contributing to the scarcity of works in this domain.

% One important aspect of the heuristics is whether they imitate the iterative nature of manual sound design---that is, listening to the output and adjusting the synthesizer parameters recursively until the desired goal is reached. Supervised methods generate their final output in a single step, while direct sound optimization methods refine their outputs iteratively until termination, which resembles manual sound design more closely.

\subsection{Iterative Sound Matching}
\label{sec:iterative_sound_matching}
 \todo{Figure~\ref{fig:sound_design_loop_iterative} shows a general model for iterative sound-matching. To begin the process, the parameters are arbitrarily initialized (usually random generation), the similarity of the target and output is measured, and the parameter is \textit{optimized} with the goal of increasing the similarity, or alternatively, reducing the loss. This process repeats until termination. 
}
\begin{figure}[ht]
    \centering
\begin{tikzpicture}[node distance=2cm, auto]

    % Nodes
    \node (start) [text centered] {\( g (\hat{\theta}) = x \)};
    \node (L) [above of=start,right of=start, text centered] {\( L(\hat{\theta},t) {\ \propto \ } \phi(x) - \phi(t) \)};
    \node (optimize) [below of=L,right of=L, text centered] {Optimize $\hat{\theta}$};
    \node (new_theta) [below of=optimize,right of = start, text centered] {New \( \hat{\theta} \)};

    % Highlight and arrow for the start node
    \draw[->, very thick, red] (start) ++(-2,1.5) -- (start)
        node[midway, below, align=center, sloped, color=red] {Start}
        node[midway, above, align=center, sloped, color=red] {Random $\hat{\theta}$};

    % Arrows with multi-line labels
    \draw[->, bend left] (start) to node[midway, right, align=center] {} (L);
    \draw[->, bend left] (L) to node[midway, right, align=center] {} (optimize);
    \draw[->, bend left] (optimize) to node[midway, below, align=center] {} (new_theta);
    \draw[->, bend left] (new_theta) to node[midway, left, align=center] {} (start);

\end{tikzpicture}
    % \caption{ Iterative approach to sound design. Synthesizer $g$ with arbitrary initialized parameters $\hat{\theta}$, creates sound $x$. The target sound $t$ is used as the goal. Parameters $\hat{\theta}$ are adjusted to minimize error (or loss) $L(\hat{\theta},t)$, where $L$ is proportional to the difference between the representations of $x$ and $t$. $\phi$ is the audio feature extractor, or representation function.}
    \caption{ Iterative approach to sound design.}
    \label{fig:sound_design_loop_iterative}
\end{figure}

In the upcoming ``Past Works'' section, we will discuss the evolution of prior sound-matching research, with \textit{differentiable iterative sound-matching} being the latest extension of the field.

\begin{table*}[t]
\centering
\caption{Summary of select works in sound-matching. Gen. refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
\setlength{\tabcolsep}{3pt} % Reduce horizontal spacing
\renewcommand{\arraystretch}{1.05} % Reduce vertical spacing
\scriptsize % Reduce font size
\resizebox{\textwidth}{!}{
\begin{tabular}{|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1cm}|>{\raggedright\arraybackslash}m{1cm}|}
\hline
\textbf{Work} & \textbf{Synthesis} & \textbf{Loss} & \textbf{Heuristics} & \textbf{Goal} & \textbf{Domain} & \textbf{Year} & \textbf{Gen. } \\
\hline
Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-differentiable FM & STFT (McAulay-Quatieri) & GA & Replicate & In & 1991 & Iter \\
\hline
Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-differentiable FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
\hline
Yee-King \textit{\textit{et al.}}~\cite{yee2018automatic} & Non-differentiable VST & P-Loss (for Supervised), Spectral (MFCC) & Supervised \& Evolutionary & Replicate & In & 2018 & Both \\
\hline
Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-differentiable VST & Spec MSS/SC P-Loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
\hline
Shier \textit{\textit{et al.}}~\cite{shier2020spiegelib} & Custom, non-diff & FFT \& Spectral & Supervised \& Direct & Replicate & In & 2020 & Both \\
\hline
Salimi \textit{\textit{et al.}}~\cite{salimi2020make} & DSP & STFT and Envelope & Supervised & Imitate & Out & 2020 & 1-shot \\
\hline
Masuda \textit{\textit{et al.}}~\cite{masuda2021soundmatch} & DDSP & P-Loss initially, fine-tune with Spec. & Supervised & Both & In \& Out & 2021 & 1-shot \\
\hline
Han \textit{\textit{et al.}}~\cite{han2023perceptual} & NN-encoder $\rightarrow$ physical model & PNP (approx. of STFT) & Supervised & Replicate & In & 2023 & 1-shot \\
\hline
Vahidi \textit{\textit{et al.}}~\cite{vahidi2023mesostructures} & Differentiable FM & JTFS/MSS & Gradient desc. & Replicate & In & 2023 & Iter \\
\hline
Barkan \textit{et al.}~\cite{barkan2023inversynthII}& Diff. Synthesizer Proxy& P-Loss + L1 Spec (weighted) &  Supervised (init.), Semi-Supervised & Replicate& In & 2023 & Both
\\
\hline
Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
\hline
Cherep \textit{et al.}~\cite{creativecherep2024} & DDSP & CLAP & Evolutionary & Imitate & Out & 2024 & Iter \\
\hline
\end{tabular}
}

\label{tab:summary}
\end{table*}

\section{Past Works}
\label{sec:related_works}
Table~\ref{tab:summary} contains an overview of relevant literature, covering important works in sound-matching. adsfEarlier works in this domain focused on FM synthesis and genetic algorithms (or direct optimization methods)\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}. Along with the exponential increase in computation power, later works utilized a wider range of approaches to parameter search, estimation, and audio synthesis; particularly with supervised learning and neural networks (\gls{NN})~\cite{yee2018automatic,engel2020ddsp,esling2019flow}.

\subsection{Earlier Works}
FM synthesizers are simple to implement, yet very expressive~\cite{chowning1973synthesis}. Perhaps the earliest foundational work in sound-matching is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds with FM synthesis using one carrier, one modulator, and the corresponding envelopes:
\[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
\]
Where $f(t)$ is the output signal, $I_c(t)$ is the carrier amplitude, $\omega_c$ is the carrier frequency, and $I_m(t)$ and $\omega_m$ are the envelope and frequency for the modulator~\cite{justice1979analytic}.

 The sound-matching works that followed used other heuristics for finding the synthesizer parameters, yet the simple FM synthesizer structure remained unchanged. Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri method~\cite{mcaulay1986speech} for measuring loss.

 % MQ calculates the STFT and marks the start and end of \textit{trends} in a sound with the assumption that the harmonic content belongs to trends that have a beginning and an end.

>>>>

\section{Experimental Setup}
\label{sec:experiment_setup}
Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers. What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and---most importantly---manual listening tests.

\textbf{We hypothesize that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}. If true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer. 
% For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).

% We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods. \textcolor{highlight}{We hypothesize that the best performing loss function is influenced by the synthesis techniques utilized in the synthesizer}. This hypothesis targets the issues of \LossSelect~and \SynthSelect.

We test four different loss functions (described in Section~\ref{sec:loss_implementation}) and compare their success in sound-matching using four different synthesizers (Section~\ref{sec:programs}). We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where Likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of final outputs and the corresponding targets. Details regarding the evaluation methods are described in Section~\ref{sec:evaluation_manual_auto}.
